Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.Rmd
Commit 2024-06-09 19:11

Main factors responsible for the attenuation of the broadband downward solar
radiation under cloud free atmospheres are aerosols and water vapor.  At
Thessaloniki, such measurements are available since 2003 from the Cimel sun
photometer that belongs to the Aerosol Robotic Network (AERONET) [@Giles2019;
@Buis1998].  The mean monthly AOD on different wavelengths are provided by
AERONET, along with the equivalent water column height in the atmosphere.

<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.

@@

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents, mainly AOD. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.

>>>>

Main factors responsible for the attenuation of the broadband downward solar
radiation under cloud free atmospheres are aerosols and water vapor.  At
Thessaloniki, such measurements are available since 2003 from the Cimel sun
photometer that belongs to the Aerosol Robotic Network (AERONET) [@Giles2019;
@Buis1998].  From the observations in the period 2003 -- 2023 we calculated the
monthly climatological means and standard deviation ($\sigma$) for the aerosol
optical depth (AOD) at $500\,\text{nm}$ and the equivalent height of the water column
(WC).  The monthly climatological values of AOD and water column, as well as
combinations with additional offsets of $\pm1\sigma$ and $\pm2\sigma$ , were used as
inputs to the RTM in the construction of the LUT.

For each measurement of the dataset, we calculated from the LUT a
$\text{GHI}_\text{ref}$ value for the respective season and SZA (by linear
interpolation), and for the climatological values of AOD and WC of the respective
month. The same procedure was followed for the estimation of the
$\text{GHI}_\text{ref}$ for all combinations of the AOD and WC with the above
mentioned standard deviation offsets.  Finally, each $\text{GHI}_\text{ref}$ value
was adjusted to the actual Sun-Earth distance derived by the Astropy software library
[@AstropyCollaboration2022].

@@

<!-- Aerosol change -->
The above clear sky reference values are based on the climatological AOD and WC,
hence they cannot describe accurately the long term variation of
$\text{GHI}_\text{ref}$ due to changes in the two atmospheric constituents, mainly
AOD.  As reported by @Natsis2023, there is a long term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data was
attributed to long-term changes in aerosol effects.

In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.

### 2024-06-09 article/article.tex
Commit 2024-06-09 19:11

Main factors responsible for the attenuation of the broadband downward
solar radiation under cloud free atmospheres are aerosols and water
vapor. At Thessaloniki, such measurements are available since 2003 from
the Cimel sun photometer that belongs to the Aerosol Robotic Network
(AERONET) \citep{Giles2019, Buis1998}. The mean monthly AOD on different
wavelengths are provided by AERONET, along with the equivalent water
column height in the atmosphere.

For the atmospheric characteristics, we iterated on combinations of AOD
at \(500\,\text{nm}\) (\(\tau_{500\text{nm}}\)) with additional offsets
of \(\pm1\) and \(\pm2\sigma\), and water column (\(w\)), also with
offsets of \(\pm1\) and \(\pm2\sigma\).

@@

The above clear sky reference values are not enough to describe the long
term variation, due to the changes of the atmospheric constituents,
mainly AOD. As observed by \citet{Natsis2023}, there is a long term
brightening effect, over Thessaloniki for the studied period. In order
to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, \citet{Kazadzis2007} have
found a change of \(-3.8\,\%\) per year at AOD on \(340\,\text{nm}\),
using a Brewer spectrophotometer. For 2005 to 2023, we derived the trend
from the available monthly climatological data of AERONET.

>>>>

Main factors responsible for the attenuation of the broadband downward
solar radiation under cloud free atmospheres are aerosols and water
vapor. At Thessaloniki, such measurements are available since 2003 from
the Cimel sun photometer that belongs to the Aerosol Robotic Network
(AERONET) \citep{Giles2019, Buis1998}. From the observations in the
period 2003 -- 2023 we calculated the monthly climatological means and
standard deviation (\(\sigma\)) for the aerosol optical depth (AOD) at
\(500\,\text{nm}\) and the equivalent height of the water column (WC).
The monthly climatological values of AOD and water column, as well as
combinations with additional offsets of \(\pm1\sigma\) and
\(\pm2\sigma\) , were used as inputs to the RTM in the construction of
the LUT.

For each measurement of the dataset, we calculated from the LUT a
\(\text{GHI}_\text{ref}\) value for the respective season and SZA (by
linear interpolation), and for the climatological values of AOD and WC
of the respective month. The same procedure was followed for the
estimation of the \(\text{GHI}_\text{ref}\) for all combinations of the
AOD and WC with the above mentioned standard deviation offsets. Finally,
each \(\text{GHI}_\text{ref}\) value was adjusted to the actual
Sun-Earth distance derived by the Astropy software library
\citep{AstropyCollaboration2022}.

@@

The above clear sky reference values are based on the climatological AOD
and WC, hence they cannot describe accurately the long term variation of
\(\text{GHI}_\text{ref}\) due to changes in the two atmospheric
constituents, mainly AOD. As reported by \citet{Natsis2023}, there is a
long term brightening effect in the GHI data of Thessaloniki for the
period 1993 -- 2023, which for clear-sky data was attributed to
long-term changes in aerosol effects. In order to better incorporate
this change, we used, estimations of AOD change derived from two
sources. For 1993 -- 2005, \citet{Kazadzis2007} have found a change of
\(-3.8\,\%\) per year at AOD on \(340\,\text{nm}\), using a Brewer
spectrophotometer. For 2005 to 2023, we derived the trend from the
available monthly climatological data of AERONET.
